Cities (Deaths & Case)

### 2020-11-13 KPMG Team/Dashboard/visualization_scripts/cities.Rmd

```{r}
```{r}
# ----------- LOAD PACKAGES ------------- #

@@

#nytimes county-level cases & deaths data from online link 
nyt <- read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv")

@@

# US City & FIPS code boundaries 
# https://simplemaps.com/data/us-cities license 
city <- read_csv("uscities.csv")

View(city)
names(city)
```

```{r}
names(city)

@@

# CITY DATASET 
data_city <- city %>% 
  rename(fips = county_fips) %>% 
  group_by(city) %>% 
  mutate(city_population = sum(population))

View(nyt)

# CITY & COVID MERGE 
# only need county to merge both datasets, will keep just city areas 
city_covid_merge <- data_city %>% 
  left_join(nyt, by = "fips") %>% 
  # select(city, state_id, state_name, fips, county_name, date, cases, deaths, population, city_population, incorporated, density, lat, lng) %>%
  select(city, state_id, state_name, fips, county_name, date, cases, deaths, population, city_population, incorporated, density, lat, lng) %>%
  filter(date >= lubridate::as_date("2020-01-24") & 
           date <= lubridate::as_date(Sys.Date() - 8))

# get 7-day mean of cases with zoo::rolling mean and dplyr::lag function 
city_covid_cases <- data.frame(city_covid_merge) %>% 
  group_by(city, date) %>% 
  dplyr::mutate(cases_sum = max(cases)) %>% # choose max cases for each date because some dates report more than once
  mutate(new_cases = cases_sum - dplyr::lag(x = cases_sum, n = 1, order_by = date)) %>%
  mutate(new_cases_mean = rollmean(new_cases, k = 7, fill = NA))

View(city_covid_cases)
# same as above, but for deaths
msa_covid_deaths <- data.frame(msa_covid_merge) %>% 
  group_by(msa, date) %>% 
  dplyr::summarise(deaths_sum = sum(deaths, na.rm = TRUE)) %>% 
  mutate(new_deaths = deaths_sum - dplyr::lag(x = deaths_sum, n = 1, order_by = date)) %>%
  mutate(new_deaths_mean = rollmean(new_deaths, k = 7, fill = NA))

# join deaths and cases numbers together (not sure if I can do it in one?)
# contains
msa_covid <- msa_covid_merge %>% 
  left_join(msa_covid_cases, by = c("msa", "date")) %>% 
  left_join(msa_covid_deaths, by = c("msa", "date")) %>% 
  arrange(desc(msa_population))

>>>>

```{r}
# ----------- LOAD PACKAGES ------------- #

@@

# nytimes county-level cases & deaths data from online link 
# a few less steps of data manipulation required, but essentially the same from JHU.
nyt <- read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv")

@@

# US City & FIPS code boundaries 
# https://simplemaps.com/data/us-cities license 
city_data <- read_csv("data/uscities.csv")

```{r}

@@

# CITY DATASET 
# cities are actually smaller than counties, e.g. New York City is part of New York County

# 
# city_data %>% 
#   data.frame() %>% 
#   arrange(desc(population)) %>% 
#   slice(1:300)
# 
# View(city_data_2)
# 
# View(city_data)
#   
# 
# city_data_2 <- city_data %>% 
#   data.frame() %>%
#   mutate(fips = county_fips) %>% 
#   # mutate(state = state_name) %>% 
#   # mutate(state_code = state_id) %>% 
#   # mutate(county = county_name) %>% 
#   # mutate(county_popn = population) %>%
#   group_by(state_name, city) %>% 
#   mutate(city_population = sum(population)) %>% 
#   select(city, state_name, state_id, city_population, fips, county_name, population)

# 
# city_state <- city_data_renamed %>% 
#   select(city, state_id, state_name) %>% 
#   distinct()
# 
# top300 <- city_data_renamed %>%
#   data.frame() %>%
#   select(city, city_population) %>%
#   arrange(desc(city_population)) %>% # %>% 
#   distinct() %>%
#   slice(1:300)
# 
# city_state_300 <- top300 %>%
#   left_join(data.frame(city_data_renamed), by = c("city", "city_population"))
# 
# View(city_state_300)
# 
# city_data_renamed
# 
# city_df <- top300
# View(top300)
# 
# 
# city_data_2 %>% 
#   arrange(desc(city_population))

View(city_data_2)
```

```{r}

# CITY & COVID MERGE 
# only need county to merge both datasets, will keep just city areas 
city_covid_merge <- city_df %>% 
  left_join(nyt, by = "fips") %>% 
  # select(city, state_id, state_name, fips, county_name, date, cases, deaths, population, city_population, incorporated, density, lat, lng) %>%
  # select(city, state_id, state_name, fips, county_name, date, cases, deaths, population, city_population, incorporated, density, lat, lng) %>%
    select(city, state_id, state_name, county_name, date, cases, deaths, population, city_population, incorporated, density, lat, lng) %>%
  filter(date >= lubridate::as_date("2020-01-24") & 
           date <= lubridate::as_date(Sys.Date() - 8))

# get 7-day mean of cases with zoo::rolling mean and dplyr::lag function 
city_covid_cases <- data.frame(city_covid_merge) %>% 
  group_by(city, city_population, date) %>% 
  summarise(cases_sum = sum(cases, na.rm = TRUE)) %>%
  mutate(new_cases = cases_sum - dplyr::lag(x = cases_sum, n = 1, order_by = date)) %>%
  mutate(new_cases_mean = rollmean(new_cases, k = 7, fill = NA))

# same as above, but for deaths
city_covid_deaths <- data.frame(city_covid_merge) %>% 
  group_by(city, city_population, date) %>% 
  summarise(deaths_sum = sum(deaths, na.rm = TRUE)) %>% 
  mutate(new_deaths = deaths_sum - dplyr::lag(x = deaths_sum, n = 1, order_by = date)) %>%
  mutate(new_deaths_mean = rollmean(new_deaths, k = 7, fill = NA))

# join deaths and cases numbers together (not sure if I can do it in one?)
city_covid <- city_covid_merge %>% 
  left_join(city_covid_cases, by = c("city", "date")) %>% 
  left_join(city_covid_deaths, by = c("city", "date")) %>% 
  select(city, state_id, state_name, city_population, 
       date, cases, new_cases, new_cases_mean, cases_sum,
       deaths, new_deaths, new_deaths_mean, deaths_sum) %>%
  arrange(desc(city_population))

city_covid_sum <- city_covid %>% 
  filter(date == max(city_covid$date)) %>% 
  select(city, state_id, state_name, city_population, 
       date, cases, new_cases, new_cases_mean, cases_sum, 
       deaths, new_deaths, new_deaths_mean, deaths_sum)

city_only <- data_city %>% 
  select(city, state_id, state_name, city_population) %>% 
  distinct()

city_covid_2 <- city_covid_cases %>%
  left_join(city_only, by = c("city", "city_population"))

View(city_covid_2)

View(data_city)
View(city_covid_merge)
View(city_covid_deaths)

names(data_city)
names(city_covid)

View(city_only)

View(city)
```

```{r}
